---
title: "Extracting location data from EPA Comment emails"
author: "Andres Garcia"
date: "5/12/2021"
output: html_document
---

**This is a work in progress and will be edited! The data scraping is a "quick and dirty" solution that has flaws and could be improved.**

# Version History

- **12 May 2021**: first version

# Summary

This is a notebook which goes through the process of extracting location data from EPA comment email pdf and txt files. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

Load packages.

```{r load, message = FALSE, warning = FALSE}
library(tidyverse)
```

# Run text scarper to extarct location data

It is assumed that the last line in a file contains the location data we are interested in.
Location text is not consistent, the last line not always contains location data. 
For example, a location can be "New York, 12084", "NY, 12084", or just 12084.
Locations can also be locations in Europe and or Canada. 
Sometimes the last line is not a location at all and it includes text such as "sent form iphone".

Because of these cases, and since most locations are XX #########, the text parser assumes that a valid location starts with a letter and ends with a number. If it does not, then a message to check a file is printed next to the possible incorrect location. 

**Note** that the `text_parser.py` script we are importing has to be in the same directory as this notebook. If you are interested in the Python code, take a look at that script on the GitHub repo.
**Make sure the output path exists.**
```{python}
import epa_location_parser
epa_location_parser.main(txtFolders = ['1. Mass 1', '1. Mass 1b'], 
                          pdfFolders = ['2. Mass 2 with attachments - FOTE'], 
                          output = 'nfs/mariculture-data/PPA_Data/EPA Public Comment/EPACommentsLocations/')
```


Now we can load the resulting CSV into R.
Let's load the csv for the  Part A pdf file and the csv with the locations for folder 1. Mass 1b.
```{r, message = FALSE}
parta_comments <- read_csv('/nfs/mariculture-data/PPA_Data/EPA Public Comment/EPACommentsLocations/Part A.pdf.csv')

mass_1b <- read_csv('/nfs/mariculture-data/PPA_Data/EPA Public Comment/EPACommentsLocations/1. Mass 1b.csv ')
```

