---
title: "EExamples for extracting location data from EPA Comment emails"
author: "Andres Garcia"
date: "5/13/2021"
output: html_document
---

# Summary

**This is notebook with examples on how to extract location data from EPA comment emails.
It covers how to extract locations from pdfs and txt files.**

# Version History

- **12 May 2021**: first version

# Setup

```{r}
library(pdftools)
library(dplyr)
```

# PDFs Example

Specify the path to the folder contianing the pdf files and read in the files. 
`pdfFiles` contains a list of pdfs to be processed.
```{r}
pathPdfs = "/nfs/mariculture-data/PPA_Data/EPA Public Comment/2. Mass 2 with attachments - FOTE"
pdfFiles <- list.files(path = pathPdfs, pattern = "pdf",  full.names = TRUE)
```

Conver pdfs to text.
Split into individual pages for each pdf.
Pages Text contains the pages for each of the pdf files.
```{r}
pagesText <- lapply(pdfFiles, pdf_text)  
```

Split pages into lines.
```{r}
pagesLines <- lapply(text, function(i) {
  strsplit(i, "\n")
})
```

Get last line containing possible location information.
```{r}
locationLines <- lapply(pagesLines, function(i) {
  tail(i[[1]], n=1)
})

# make into a dataframe for easier reading
locationdf <- as.data.frame(locationLines) %>% gather()
```

Add flag if location is not valid. 
Notice for pdf files the index is also the page number. 
Check the page to verify a location if flag is present.
```{r}
locationdf$flag <- lapply(locationdf$value, function(i) {
  case_when(
    # check 1st two character are letters
    grepl("[a-zA-Z]", substring(i, 1, 2)) == FALSE |
    # check last 9 are digits
    grepl("\\d", substr(i, nchar(i) - 9 + 1,
                        nchar(i))) == FALSE ~ "Check"
  )
})

# keep columns of interest, this can be stored as a csv file
locationdf <- locationdf[,2:3]
View(locationdf)
```

# Text Files Example

Specify the path to the txt files folder. 
```{r}
pathMass1 = "/nfs/mariculture-data/PPA_Data/EPA Public Comment/1. Mass 1"
mass1TxtFiles <- list.files(path = pathMass1, pattern = "txt",  full.names = TRUE)
```

Read txt files.
```{r}
# we will just work with the first 10 txt files in the folder
# remove [1:10] to process all files
textFiles <- lapply(mass1TxtFiles[1:10], function(i) {
  read.delim(i)
})
```

Get last lines containing possible location information.
```{r}
locationLines <- lapply(textFiles, function(i) {
  tail(i[[1]], n=1)
})

# make into a dataframe for easier reading
locationdf <- as.data.frame(locationLines) %>% gather()
```

Add flag if location is not valid.
Not valid = 1st two character are not letter or last 9 characters are not numbers.
```{r}
locationdf$flag <- lapply(locationdf$value, function(i) {
  case_when(
    # check if 1st two character are letters
    grepl("[a-zA-Z]", substring(i, 1, 2)) == FALSE |
    # check if last 9 are digits
    grepl("\\d", substr(i, nchar(i) - 9 + 1,
                        nchar(i))) == FALSE ~ "Check"
  )
})

# keep columns of interest, this can be stored as a csv file
locationdf <- locationdf[,2:3]
View(locationdf)
```

